Figure 7 - heterodimer

### 2024-08-03 ms_figure_7_heterodimer.Rmd
Clean up Fig 7 code, add panel label

```{r}
## Read data with distances from nuclear envelope + cell radius and 
## calculate relative distance from NE
dist_ne <- read.delim(params$dist_from_ne) |>
    group_by(Condition, Replicate) |>
    mutate(rel_distance_from_ne = Distance_from_nuc_envelope / Radius_of_cell) |>
    mutate(Condition = ifelse(Condition == "ntu1D", "ntu1\U141E", 
                              ifelse(Condition == "ntu2D", "ntu2\U141E", Condition))) |>
    mutate(SampleID = factor(paste0(Condition, "\n", Replicate),
                             levels = c("wt\nrep1", "wt\nrep2", "ntu1\U141E\nrep1",
                                        "ntu1\U141E\nrep2", "ntu2\U141E\nrep1",
                                        "ntu2\U141E\nrep2")))

@@

## Extract t-statistics for the two comparisons
rdhigh <- rowData(sce500)
contrasts_high <- c(
    "Ntu1_500_plate_vs_compl_Ntu1_500_plate.t", 
    "Ntu2_500_plate_vs_compl_Ntu2_500_plate.t")
high <- as.data.frame(rdhigh[, c("einprotId", contrasts_high)]) |>
    setNames(c("einprotId", c("Ntu1", "Ntu2")))

@@

```{r}
#| label: rnaseq-corrplot

## Correlation between log2 fold changes for ntu1/ntu2
stopifnot(all(rownames(resList$ntu1) == rownames(resList$ntu2)))
dfcorr <- data.frame(gene = rownames(resList$ntu1), 
                     ntu1 = resList$ntu1$log2FoldChange,
                     ntu2 = resList$ntu2$log2FoldChange)
(corrplot <- list(
    ggplot(dfcorr, aes(x = ntu1, y = ntu2)) + 
        geom_abline(slope = 1, intercept = 0, linetype = "dashed", 
                    color = "grey50") + 
        geom_point(alpha = 0.25) + 
        geom_text_repel(data = dfcorr |>
                            dplyr::filter(gene %in% c("tna1", "SPCC576.17c")), 
                        aes(label = gene), 
                        min.segment.length = 0, max.overlaps = Inf) + 
        labs(x = "*ntu1\U141E* log2(fold-change)",
             y = "*ntu2\U141E* log2(fold-change)") + 
        theme_cowplot() + 
        theme(axis.title = element_markdown())
))
```

@@

resListToPlot <- lapply(resList, function(rl) {
    rl |> dplyr::filter(seqnames == chrom & 
                            start > startpos - windowsize & 
                            end < endpos + windowsize)
})
plotList <- lapply(names(resListToPlot), function(nm) {
    ggplot(resListToPlot[[nm]], 
           aes(x = start, y = log2FoldChange, 
               color = abs(log2FoldChange) > 3 & padj < 0.001)) + 
        geom_rect(aes(xmin = endpos, xmax = startpos, ymin = -Inf, ymax = Inf),
                  color = lighten(main_colors[4], amount = 0.6), fill = NA,
                  alpha = 0.1, show.legend = FALSE) + 
        geom_point(alpha = 0.6, size = 2) + 
        scale_color_manual(values = c(`FALSE` = "grey50", 
                                      `TRUE` = main_colors[4])) + 
        geom_text_repel(aes(label = ifelse(abs(log2FoldChange) > 3 & padj < 0.001,
                                           as.character(gene), "")),
                        size = 2.5, max.overlaps = 30, force = 10) + 
        scale_x_continuous(labels = scales::label_number(scale = 1e-6)) + 
        scale_y_continuous(limits = c(-7, 7)) + 
        labs(x = "Genome position (Mb)", y = "log2(fold-change)",
             title = paste0(nm, "\U141E")) + 
        theme_cowplot() + 
        theme(legend.position = "none",
              plot.title = element_text(face = "italic"))
})
genomeplot <- cowplot::plot_grid(plotlist = plotList, ncol = 1)
```

@@

## Create data frames for plotting
plotdf_chip1 <- covs_chip |>
    mutate(sample = sub("_rep[0-9]", "", sample)) |>
    mutate(exptype = stringr::str_extract(sample, "ChIP|IN")) |>
    mutate(pair = sub("_ChIP|_IN", "", sample)) |>
    mutate(flag = .capitalize(sub("_*TAG", "", sub(".*_KO_", "", pair)))) |>
    mutate(label = ifelse(pair == "ntu1_KO_ntu2TAG", "*ntu1\U141E*",
                          ifelse(pair == "ntu1_TAG", "*ntu2\U207A*",
                                 ifelse(pair == "ntu2_TAG", "*ntu1\U207A*",
                                        ifelse(pair == "ntu2_KO_ntu1TAG", "*ntu2\U141E*", 
                                               ""))))) |>
    mutate(cpm = score) |>
    mutate(pos = start) |>
    group_by(flag, seqnames) |>
    mutate(cpmnorm = cpm / max(abs(cpm))) |>
    ungroup() |>
    filter(pair %in% c("ntu2_TAG", "ntu2_KO_ntu1TAG", 
                       "ntu1_TAG", "ntu1_KO_ntu2TAG")) |>
    mutate(pair = factor(pair, levels = c("ntu1_TAG", "ntu2_KO_ntu1TAG",
                                          "ntu2_TAG", "ntu1_KO_ntu2TAG")))
plotdf_rna <- covs_rna |>
    filter(sample %in% c("SPBC530.08_KO_mRNA_rep1",
                         "SPBC16G5.16_KO_mRNA_rep1",
                         "wt_KO_mRNA_rep1")) |>
    mutate(sample = sub("_KO_mRNA_rep[0-9]", "", sample)) |>
    mutate(sample = replace(sample, sample == "SPBC16G5.16", "ntu1\U141E")) |>
    mutate(sample = replace(sample, sample == "SPBC530.08", "ntu2\U141E")) |>
    mutate(gene = sample) |>
    group_by(seqnames) |>
    mutate(cpmnorm = cpm / max(abs(cpm))) |>
    ungroup() |>
    mutate(sample = factor(sample, 
                           levels = c("wt", "ntu1\U141E", "ntu2\U141E")))
plotdf_label <- covs_rna |>
    filter(sample == "wt_KO_mRNA_rep1") |>
    mutate(cpmnorm = 0)

@@

    # geom_text(aes(x = (start2 + end2) / 2, y = y - 0.5), 
    #           size.unit = "in", size = 0.1) + 

@@

    g1c + theme_void() + theme(strip.text = element_blank()),
    g1a + theme_void() + theme(legend.position = "none", 

@@

    g1b + theme_void() + theme(strip.text.y = element_text(size = 12, 

@@

    g1c + theme_bw() + theme(axis.text.x = element_text(size = 5)),
    g1a + theme_bw() + theme(legend.position = "none",

@@

    g1b + theme_bw() + theme(legend.position = "none",

@@

cowplot::plot_grid(
    cowplot::plot_grid(
        gg_ipdimer, 
        fig7c,
        nrow = 1, 
        rel_widths = c(1, 2), 
        labels = c("A", "C"),
        align = "h", axis = "bt"
    ),
    cowplot::plot_grid(
        fig7b, 
        gcov,
        rel_widths = c(1, 1),
        labels = c("B", "D")
    ),
    cowplot::plot_grid(
        fig7e,
        NULL,
        fig7f,
        NULL,
        p_zone + theme(legend.position = "top"), 
        labels = c("E", "", "F", "", "G"),
        # rel_widths = c(7.65, 0.5, 1.45, 0.5, 9)
        rel_widths = c(6, 0.75, 1.2, 0.75, 10.4), nrow = 1
    ), 
    ncol = 1
)
```

@@

plotdf_chip2 <- covs_chip |>
    mutate(sample = sub("_rep[0-9]", "", sample)) |>
    mutate(exptype = stringr::str_extract(sample, "ChIP|IN")) |>
    mutate(pair = sub("_ChIP|_IN", "", sample)) |>
    mutate(flag = sub("_*TAG", "", sub(".*_KO_", "", pair))) |>
    mutate(label = ifelse(pair == "ntu2_delTMD", "delTMD",
                          ifelse(pair == "ntu2_TAG", "ntu2+", ""))) |>
    mutate(cpm = score) |>
    mutate(pos = start) |>
    group_by(seqnames) |>
    mutate(cpmnorm = cpm / max(abs(cpm))) |>
    ungroup() |>
    filter(pair %in% c("ntu2_TAG", "ntu2_delTMD")) |>
    mutate(pair = ifelse(pair == "ntu2_delTMD", "Ntu2\U141E\U1D40\U1D39\U1D30", 
                         ifelse(pair == "ntu2_TAG", "Ntu2", pair))) |>
    mutate(pair = factor(pair, levels = c("Ntu2", "Ntu2\U141E\U1D40\U1D39\U1D30")))

@@

figs6a <- ggdraw() + 
    draw_image("schematics/FigS7A_Nattou_AF2_fl_Arial.png")

cowplot::plot_grid(
    figs6a,
    cowplot::plot_grid(
        p_rel_dis,
        genomeplot_tpm,
        nrow = 1, 
        rel_widths = c(1, 1), 
        labels = c("B", "C")
    ),
    cowplot::plot_grid(
        genomeplot,
        gcovdel,
        # maplots$ID83delTMD,
        nrow = 1, 
        rel_widths = c(1, 1),
        labels = c("D", "E")
    ),
    ncol = 1, rel_heights = c(1, 1, 1)
)
```

>>>>

```{r}
## Read data with distances from nuclear envelope + cell radius and 
## calculate relative distance from NE
dist_ne <- read.delim(params$dist_from_ne) |>
    group_by(Condition, Replicate) |>
    mutate(rel_distance_from_ne = Distance_from_nuc_envelope / Radius_of_cell) |>
    mutate(Condition = ifelse(Condition == "ntu1D", "ntu1\U141E", 
                              ifelse(Condition == "ntu2D", "ntu2\U141E", Condition))) |>
    mutate(SampleID = factor(paste0(Condition, "\n", Replicate),
                             levels = c("wt\nrep1", "wt\nrep2", "ntu1\U141E\nrep1",
                                        "ntu1\U141E\nrep2", "ntu2\U141E\nrep1",
                                        "ntu2\U141E\nrep2"))) |>
    dplyr::ungroup()

@@

## Extract t-statistics for the two comparisons
rdhigh <- rowData(sce500)
contrasts_high <- c(
    "Ntu1_500_plate_vs_compl_Ntu1_500_plate.t", 
    "Ntu2_500_plate_vs_compl_Ntu2_500_plate.t")
high <- as.data.frame(rdhigh[, c("einprotId", contrasts_high)]) |>
    setNames(c("einprotId", "Ntu1", "Ntu2"))

@@

```{r}
#| label: rnaseq-corrplot
#| warning: false

## Correlation between log2 fold changes for ntu1/ntu2
stopifnot(all(rownames(resList$ntu1) == rownames(resList$ntu2)))
dfcorr <- data.frame(gene = rownames(resList$ntu1), 
                     ntu1 = resList$ntu1$log2FoldChange,
                     ntu2 = resList$ntu2$log2FoldChange)
(corrplot <- list(
    ggplot(dfcorr, aes(x = ntu1, y = ntu2)) + 
        geom_abline(slope = 1, intercept = 0, linetype = "dashed", 
                    color = na_color) + 
        geom_point(alpha = 0.25) + 
        geom_text_repel(data = dfcorr |>
                            dplyr::filter(gene %in% c("tna1", "SPCC576.17c")), 
                        aes(label = gene), 
                        min.segment.length = 0, max.overlaps = Inf) + 
        labs(x = "*ntu1\U141E* log2(fold-change)",
             y = "*ntu2\U141E* log2(fold-change)") + 
        theme_cowplot() + 
        theme(axis.title = element_markdown())
))
```

@@

resListToPlot <- lapply(resList, function(rl) {
    rl |> dplyr::filter(seqnames == chrom & 
                            start > startpos - windowsize & 
                            end < endpos + windowsize)
})
plotList <- lapply(names(resListToPlot), function(nm) {
    ggplot(resListToPlot[[nm]], 
           aes(x = start, y = log2FoldChange, 
               color = abs(log2FoldChange) > 3 & padj < 0.001)) + 
        geom_rect(aes(xmin = endpos, xmax = startpos, ymin = -Inf, ymax = Inf),
                  color = lighten(main_colors[4], amount = 0.6), fill = NA,
                  alpha = 0.1, show.legend = FALSE) + 
        geom_point(alpha = 0.6, size = 2) + 
        scale_color_manual(values = c(`FALSE` = na_color, 
                                      `TRUE` = main_colors[4])) + 
        geom_text_repel(aes(label = ifelse(abs(log2FoldChange) > 3 & padj < 0.001,
                                           as.character(gene), "")),
                        size = 2.5, max.overlaps = 30, force = 10) + 
        scale_x_continuous(labels = scales::label_number(scale = 1e-6)) + 
        scale_y_continuous(limits = c(-7, 7)) + 
        labs(x = "Genome position (Mb)", y = "log2(fold-change)",
             title = paste0(nm, "\U141E")) + 
        theme_cowplot() + 
        theme(legend.position = "none",
              plot.title = element_text(face = "italic"))
})
genomeplot <- cowplot::plot_grid(plotlist = plotList, ncol = 1)
```

@@

## Create data frames for plotting
plotdf_chip1 <- covs_chip |>
    mutate(sample = sub("_rep[0-9]", "", sample)) |>
    mutate(exptype = stringr::str_extract(sample, "ChIP|IN")) |>
    mutate(pair = sub("_ChIP|_IN", "", sample)) |>
    mutate(flag = .capitalize(sub("_*TAG", "", sub(".*_KO_", "", pair)))) |>
    mutate(label = ifelse(pair == "ntu1_KO_ntu2TAG", "*ntu1\U141E*",
                          ifelse(pair == "ntu1_TAG", "*ntu2\U207A*",
                                 ifelse(pair == "ntu2_TAG", "*ntu1\U207A*",
                                        ifelse(pair == "ntu2_KO_ntu1TAG", "*ntu2\U141E*", 
                                               ""))))) |>
    mutate(cpm = score) |>
    mutate(pos = start) |>
    group_by(flag, seqnames) |>
    mutate(cpmnorm = cpm / max(abs(cpm))) |>
    ungroup() |>
    filter(pair %in% c("ntu2_TAG", "ntu2_KO_ntu1TAG", 
                       "ntu1_TAG", "ntu1_KO_ntu2TAG")) |>
    mutate(pair = factor(pair, levels = c("ntu1_TAG", "ntu2_KO_ntu1TAG",
                                          "ntu2_TAG", "ntu1_KO_ntu2TAG")))
table(plotdf_chip1$sample, plotdf_chip1$flag)
table(plotdf_chip1$sample, plotdf_chip1$label)
plotdf_rna <- covs_rna |>
    filter(sample %in% c("SPBC530.08_KO_mRNA_rep1",
                         "SPBC16G5.16_KO_mRNA_rep1",
                         "wt_KO_mRNA_rep1")) |>
    mutate(sample = sub("_KO_mRNA_rep[0-9]", "", sample)) |>
    mutate(sample = replace(sample, sample == "SPBC16G5.16", "ntu1\U141E")) |>
    mutate(sample = replace(sample, sample == "SPBC530.08", "ntu2\U141E")) |>
    mutate(gene = sample) |>
    group_by(seqnames) |>
    mutate(cpmnorm = cpm / max(abs(cpm))) |>
    ungroup() |>
    mutate(sample = factor(sample, 
                           levels = c("wt", "ntu1\U141E", "ntu2\U141E")))
plotdf_label <- covs_rna |>
    filter(sample == "wt_KO_mRNA_rep1") |>
    mutate(cpmnorm = 0)

@@

@@

    g1c + theme_void() + 
        theme(strip.text = element_blank()),
    g1a + theme_void() + 

@@

    g1b + theme_void() + 
        theme(strip.text.y = element_text(size = 12, 

@@

    g1c + theme_bw() + 
        theme(axis.text.x = element_text(size = 5)),
    g1a + theme_bw() + 

@@

    g1b + theme_bw() + 

@@

cowplot::plot_grid(
    cowplot::plot_grid(
        gg_ipdimer, 
        fig7c,
        nrow = 1, 
        rel_widths = c(1, 2), 
        labels = c("A", "C"),
        align = "h", axis = "bt"
    ),
    cowplot::plot_grid(
        fig7b, 
        gcov,
        rel_widths = c(1, 1),
        labels = c("B", "D")
    ),
    cowplot::plot_grid(
        fig7e,
        NULL,
        fig7f,
        NULL,
        p_zone + theme(legend.position = "top"), 
        labels = c("E", "", "F", "", "G"),
        rel_widths = c(6, 0.75, 1.2, 0.75, 10.4), nrow = 1
    ), 
    ncol = 1
)
```

@@

plotdf_chip2 <- covs_chip |>
    mutate(sample = sub("_rep[0-9]", "", sample)) |>
    mutate(exptype = stringr::str_extract(sample, "ChIP|IN")) |>
    mutate(pair = sub("_ChIP|_IN", "", sample)) |>
    mutate(flag = sub("_*TAG", "", sub(".*_KO_", "", pair))) |>
    mutate(label = ifelse(pair == "ntu2_delTMD", "delTMD",
                          ifelse(pair == "ntu2_TAG", "ntu2+", ""))) |>
    mutate(cpm = score) |>
    mutate(pos = start) |>
    group_by(seqnames) |>
    mutate(cpmnorm = cpm / max(abs(cpm))) |>
    ungroup() |>
    filter(pair %in% c("ntu2_TAG", "ntu2_delTMD")) |>
    mutate(pair = ifelse(pair == "ntu2_delTMD", "Ntu2\U141E\U1D40\U1D39\U1D30", 
                         ifelse(pair == "ntu2_TAG", "Ntu2", pair))) |>
    mutate(pair = factor(pair, levels = c("Ntu2", "Ntu2\U141E\U1D40\U1D39\U1D30")))
table(plotdf_chip2$sample, plotdf_chip2$flag)
table(plotdf_chip2$sample, plotdf_chip2$pair)

@@

figs7a <- ggdraw() + 
    draw_image("schematics/FigS7A_Nattou_AF2_fl_Arial.png")

cowplot::plot_grid(
    figs7a,
    cowplot::plot_grid(
        p_rel_dis,
        genomeplot_tpm,
        nrow = 1, 
        rel_widths = c(1, 1), 
        labels = c("B", "C")
    ),
    cowplot::plot_grid(
        genomeplot,
        gcovdel,
        nrow = 1, 
        rel_widths = c(1, 1),
        labels = c("D", "E")
    ),
    labels = c("A", "", ""), 
    ncol = 1, rel_heights = c(1, 1, 1)
)
```
